Within and beyond: chert procurement patterns during the Upper Palaeolithic in Southwesternmost Iberia

### 2025-05-02 appendix1.qmd
Macbook session. Tests to latex chapter structure 01

## XRD results

In this subsection we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC. The original data necessary to produce these visualisations can be found and freely accessed in our online compendium, in the associated GitHub project under the path "../data/xrd_data".

```{r}
#| label: somxrd-library

library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

```

```{r}
#| label: somxrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

xrd_marks <- xrd_marks |> # Correct the quartz denomination on base dataset
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# Read individual xrd files
sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
sp18_xrd <- read_csv('data/xrd_data/SP18_xrd.csv')
sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')

```

```{r}
#| label: fig-xrd-sp6
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP6_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

# SP6 Cabo S. Vicente
sp6_marks <- xrd_marks |>
  filter(Sample == "SP6")
sp6_perc <- xrd_perc |>
  filter(Sample == "SP6") |>
  select(-Sample, -ESD)

sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))

```

```{r}
#| label: fig-xrd-sp7
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP7_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp7_marks <- xrd_marks |>
  filter(Sample == "SP7")
sp7_perc <- xrd_perc |>
  filter(Sample == "SP7") |>
  select(-Sample, -ESD)

sp7_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = `Quartz (SiO2)`) |> 
  rename(Dolomite = `Dolomite (CaMg(CO3)2)`) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp7_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp7_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp7_perc)))

```

```{r}
#| label: fig-xrd-sp10
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP10_PdA). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp10_marks <- xrd_marks |>
  filter(Sample == "SP10")
sp10_perc <- xrd_perc |>
  filter(Sample == "SP10") |>
  select(-Sample, -ESD)

sp10_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp10_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp10_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp10_perc)))

```

```{r}
#| label: fig-xrd-sp14
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP14_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp14_marks <- xrd_marks |>
  filter(Sample == "SP14")
sp14_perc <- xrd_perc |>
  filter(Sample == "SP14") |>
  select(-Sample, -ESD)

sp14_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp14_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp14_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp14_perc)))

```

```{r}
#| label: fig-xrd-sp18
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp18_marks <- xrd_marks |>
  filter(Sample == "SP18")
sp18_perc <- xrd_perc |>
  filter(Sample == "SP18") |>
  select(-Sample, -ESD)

sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))

```

```{r}
#| label: fig-xrd-sp21
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Belixe Sul, north of Praia do Belixe (SP21_BLS). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp21_marks <- xrd_marks |>
  filter(Sample == "SP21")
sp21_perc <- xrd_perc |>
  filter(Sample == "SP21") |>
  select(-Sample, -ESD)

sp21_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp21_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp21_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp21_perc)))

```

```{r}
#| label: fig-xrd-sp28
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Aspa, north of Cabo S. Vicente (SP28_ASP). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp28_marks <- xrd_marks |>
  filter(Sample == "SP28")
sp28_perc <- xrd_perc |>
  filter(Sample == "SP28") |>
  select(-Sample, -ESD)

sp28_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp28_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp28_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp28_perc)))

```

```{r}
#| label: fig-xrd-sp30
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP30_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp30_marks <- xrd_marks |>
  filter(Sample == "SP30")
sp30_perc <- xrd_perc |>
  filter(Sample == "SP30") |>
  select(-Sample, -ESD)

sp30_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp30_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp30_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp30_perc)))

```

```{r}
#| label: fig-xrd-sp33
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP33_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp33_marks <- xrd_marks |>
  filter(Sample == "SP33")
sp33_perc <- xrd_perc |>
  filter(Sample == "SP33") |>
  select(-Sample, -ESD)

sp33_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp33_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp33_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp33_perc)))

```

```{r}
#| label: fig-xrd-sp34
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP34_PdA_a). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp34_marks <- xrd_marks |>
  filter(Sample == "SP34")
sp34_perc <- xrd_perc |>
  filter(Sample == "SP34") |>
  select(-Sample, -ESD)

sp34_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp34_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp34_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp34_perc)))

```

```{r}
#| label: fig-xrd-sp40
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ferrel (SP40_b). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp40_marks <- xrd_marks |>
  filter(Sample == "SP40")
sp40_perc <- xrd_perc |>
  filter(Sample == "SP40") |>
  select(-Sample, -ESD)

sp40_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = SiO2) |> 
  select(-SiO2_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp40_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp40_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp40_perc)))

```

```{r}
#| label: fig-xrd-sp50
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP50_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp50_marks <- xrd_marks |>
  filter(Sample == "SP50")
sp50_perc <- xrd_perc |>
  filter(Sample == "SP50") |>
  select(-Sample, -ESD)

sp50_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp50_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp50_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp50_perc)))

```

```{r}
#| label: fig-xrd-sp52
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP52_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp52_marks <- xrd_marks |>
  filter(Sample == "SP52")
sp52_perc <- xrd_perc |>
  filter(Sample == "SP52") |>
  select(-Sample, -ESD)

sp52_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite ,Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp52_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp52_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp52_perc)))

```

```{r}
#| label: fig-xrd-sp57
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formations) chert sample from the outcrop Jordana (SP57_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp57_marks <- xrd_marks |>
  filter(Sample == "SP57")
sp57_perc <- xrd_perc |>
  filter(Sample == "SP57") |>
  select(-Sample, -ESD)

sp57_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp57_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp57_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp57_perc)))

```

```{r}
#| label: fig-xrd-sp59
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP59_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp59_marks <- xrd_marks |>
  filter(Sample == "SP59")
sp59_perc <- xrd_perc |>
  filter(Sample == "SP59") |>
  select(-Sample, -ESD)

sp59_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp59_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp59_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp59_perc)))

```

```{r}
#| label: fig-xrd-sp61
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the secondary outcrop Peral (SP61_PER). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp61_marks <- xrd_marks |>
  filter(Sample == "SP61")
sp61_perc <- xrd_perc |>
  filter(Sample == "SP61") |>
  select(-Sample, -ESD)

sp61_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp61_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp61_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp61_perc)))

```

```{r}
#| label: fig-xrd-sp63
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Oliveiras (SP63_OLV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp63_marks <- xrd_marks |>
  filter(Sample == "SP63")
sp63_perc <- xrd_perc |>
  filter(Sample == "SP63") |>
  select(-Sample, -ESD)

sp63_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp63_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp63_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp63_perc)))

```

```{r}
#| label: fig-xrd-sp66
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Malhão (SP66_MALH). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp66_marks <- xrd_marks |>
  filter(Sample == "SP66")
sp66_perc <- xrd_perc |>
  filter(Sample == "SP66") |>
  select(-Sample, -ESD)

sp66_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp66_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp66_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp66_perc)))

```

>>>>

##

### 2025-05-02 article2.qmd
Macbook session. Tests to latex chapter structure 01

**Authors**: Joana Belmiro, Xavier Terradas, Salvador Dominguez-Bella, João Cascalheira

\newpage

@@

\newpage

>>>>

**Authors**: Joana Belmiro^1^, Xavier Terradas^2^, Salvador Dominguez-Bella^3^, João Cascalheira^1^

^1^ Interdisciplinary Center for Archaeology and the Evolution of Human Behaviour (ICArEHB), University of Algarve; ^2^ CSIC; ^3^ UCA

**Reference**: xxx

@@
